A Dependently Typed Multi-Stage Calculus \ifnotanonymous\thanks{Supported by organization x.}\fi

### 2019-06-14 text/informal-overview.tex
type variable -> type-level constant
epsilon -> varepsilon.

The subscript $\alpha$ in $\TB_\alpha M$ is a \textit{transition variable} and
a sequence of transition variables is called \textit{transition}.
The empty transition is represented by $\epsilon$.
They are used to show the thickness of brackets.
For example, $\TB_\alpha (\lambda x:\I.x+10)$ is the fragment of code which becomes $(\lambda x:\I.x+10)$ after it was run once and
$\TB_\alpha \TB_\beta (\lambda x:\I.x+10)$, which is abbreviated as $\TB_{\alpha\beta} (\lambda x:\I.x+10)$,
becomes $(\lambda x:\I.x+10)$ after it was run it twice.
\red{run twiceは表現が微妙}

@@

Another important rule about a transition variable is 
that symbols with transition variables disappear 
when the empty transition $\epsilon$ is substituted to the transition variable.
The purpose of this rule is express \verb|run| without any special symbol.
In multi-stage calculus, \verb|run| is a very important operator which changes quoted code to the original code.
For example, $(\TB_\epsilon (\lambda x:\I.x))$ is equivalent to $(\lambda x:\I.x)$.
Therefore, $\Lambda\alpha.(\TB_\alpha (\lambda x:\I.x))\ \epsilon$ becomes $\lambda x:\I.x$
In \LTP, \verb|run| is realized with application to the empty transition $\epsilon$.

@@

A type judgement of \LTP is of the form $\G \vdash M : \tau @ A$.
A transition in a judgement represents where the judgement is true.
For example, $\G \vdash (\lambda x:\I.x) : \I \to \I @ \alpha$ means 
term $\lambda x:\I.x$ has type $\I \to \I$ at transition $\alpha$.
Especially, terms without quoting exist at the empty transition $\epsilon$.
For example, $(\lambda x:\I.x)\ (1+2)$ is at $\epsilon$ transition and 
$\TB_\alpha (\lambda x:\I.x)$ is at $\alpha$ transition.
Terefore, transitions appear in typing rules, too.
\begin{center}
	\infrule{\G\vdash M:\tau @{A\alpha}}{\G\vdash \TB_{\alpha}M:\TW_{\alpha}\tau @A}{\TTB} \andalso
	\infrule{\G\vdash M:\TW_{\alpha}\tau @A}{\G\vdash \TBL_{\alpha}M:\tau @{A\alpha}}{\TTBL}
\end{center}
\TTB, corresponding to brackets, means 
if $M$ is typed $\tau$ at transition $A\alpha$ then $\TB_{\alpha}M$, quoted $M$, is typed $\TW_{\alpha}\tau$ at $A$.
\TTBL\ is converse of \TTB.

@@

First, the way of handling of constants and type-level constants is the difference between \LMD and \LTP.
We adopt a signature $\Sigma$ to handle constants and type-level constants.
This is because a signature simplifies kinding rules relating to type variables.
A signature $\Sigma$ is composed of pairs of a base type and its kind or a constant and its type.
For example, if you want to use boolean, $\Sigma = \B::*, \text{true}:\B, \text{false}:\B$
\red{具体的な導出例を出したほうがよいか? また、なぜsimpleになるのを書くべきか?}

>>>>

The subscript $\alpha$ in $\TB_\alpha M$ is a \textit{transition variable} and
a sequence of transition variables is called \textit{transition}.
The empty transition is represented by $\varepsilon$.
They are used to show the thickness of brackets.
For example, $\TB_\alpha (\lambda x:\I.x+10)$ is the fragment of code which becomes $(\lambda x:\I.x+10)$ after it was run once and
$\TB_\alpha \TB_\beta (\lambda x:\I.x+10)$, which is abbreviated as $\TB_{\alpha\beta} (\lambda x:\I.x+10)$,
becomes $(\lambda x:\I.x+10)$ after it was run it twice.
\red{run twiceは表現が微妙}

@@

Another important rule about a transition variable is 
that symbols with transition variables disappear 
when the empty transition $\varepsilon$ is substituted to the transition variable.
The purpose of this rule is express \verb|run| without any special symbol.
In multi-stage calculus, \verb|run| is a very important operator which changes quoted code to the original code.
For example, $(\TB_\varepsilon (\lambda x:\I.x))$ is equivalent to $(\lambda x:\I.x)$.
Therefore, $\Lambda\alpha.(\TB_\alpha (\lambda x:\I.x))\ \varepsilon$ becomes $\lambda x:\I.x$
In \LTP, \verb|run| is realized with application to the empty transition $\varepsilon$.

@@

A type judgement of \LTP is of the form $\G \vdash M : \tau @ A$.
A transition in a judgement represents where the judgement is true.
For example, $\G \vdash (\lambda x:\I.x) : \I \to \I @ \alpha$ means 
term $\lambda x:\I.x$ has type $\I \to \I$ at transition $\alpha$.
Especially, terms without quoting exist at the empty transition $\varepsilon$.
For example, $(\lambda x:\I.x)\ (1+2)$ is at $\varepsilon$ transition and 
$\TB_\alpha (\lambda x:\I.x)$ is at $\alpha$ transition.
Terefore, transitions appear in typing rules, too.
\begin{center}
	\infrule{\G\vdash M:\tau @{A\alpha}}{\G\vdash \TB_{\alpha}M:\TW_{\alpha}\tau @A}{\TTB} \andalso
	\infrule{\G\vdash M:\TW_{\alpha}\tau @A}{\G\vdash \TBL_{\alpha}M:\tau @{A\alpha}}{\TTBL}
\end{center}
\TTB, corresponding to brackets, means 
if $M$ is typed $\tau$ at transition $A\alpha$ then $\TB_{\alpha}M$, quoted $M$, is typed $\TW_{\alpha}\tau$ at $A$.
\TTBL\ is converse of \TTB.

@@

First, the way of handling of constants and type-level constants is the difference between \LMD and \LTP.
We adopt a signature $\Sigma$ to handle constants and type-level constants.
This is because a signature simplifies kinding rules relating to type-level constants.
A signature $\Sigma$ is composed of pairs of a base type and its kind or a constant and its type.
For example, if you want to use boolean, $\Sigma = \B::*, \text{true}:\B, \text{false}:\B$
\red{具体的な導出例を出したほうがよいか? また、なぜsimpleになるのを書くべきか?}

### 2019-06-14 text/properties.tex
分割コンパイルが失敗する
epsilon -> varepsilon.

Substitution Lemma in \LMD\ is little more complicated than an ordinary one 
because there are six types of judgment and two types of substitution in \LMD.
Substitution Lemma on Terms states that term substitution $[z \mapsto P]$ preserves
typing, kinding, well-formed kinding, term equivalence, type equivalence, and king equivalence.
Substitution Lemma on Terms states the same for stage substitution $[\beta\mapsto\epsilon]$.

@@

\begin{theorem}[Substitution Lemma on Stages]
	\begin{flalign*}
		\text{If\ } \G \V M:\tau @A
		&\text{\ then\ } \G[\beta \mapsto \epsilon]\V M[\beta \mapsto \epsilon]:\tau[\beta \mapsto \epsilon] @A[\beta \mapsto \epsilon].&\\
		\text{If\ } \G \V \tau::K @A
		&\text{\ then\ } \G[\beta \mapsto \epsilon]\V \tau[\beta \mapsto \epsilon]::K[\beta \mapsto \epsilon] @A[\beta \mapsto \epsilon].&\\
		\text{If\ } \G \V K\iskind @A
		&\text{\ then\ } \G[\beta \mapsto \epsilon]\V K[\beta \mapsto \epsilon] \iskind @A[\beta \mapsto \epsilon].&\\
		\text{If\ } \G \V M\E N : \tau @A
		&\text{\ then\ } \G[\beta \mapsto \epsilon]\V M[\beta \mapsto \epsilon]\E N[\beta \mapsto \epsilon] : \tau[\beta \mapsto \epsilon]  @A[\beta \mapsto \epsilon].&\\
		\text{If\ } \G \V \tau\E \sigma : K @A
		&\text{\ then\ } \G[\beta \mapsto \epsilon]\V \tau[\beta \mapsto \epsilon]\E \sigma[\beta \mapsto \epsilon] : K[\beta \mapsto \epsilon] @A[\beta \mapsto \epsilon].&\\
		\text{If\ } \G \V K\E J @A
		&\text{\ then\ } \G[\beta \mapsto \epsilon]\V K[\beta \mapsto \epsilon]\E J[\beta \mapsto \epsilon] @A[\beta \mapsto \epsilon].&
	\end{flalign*}
\end{theorem}

@@

\begin{theorem}[Unique Decomposition]
	If $x:\tau @\epsilon \notin \G$ and $\G \V M : \tau @ A$ then either
	\begin{enumerate}
		\item $ M \in V^A$, or
		\item there exist $B, E^A_B$, and $R^B$ such that $M = E^A_B[R^B]$ with $B = \epsilon$ or $B = \beta$ for some $\beta$.
	\end{enumerate}
\end{theorem}

@@

\begin{theorem}[Progress]
	If $x:\tau @\epsilon \notin \G$ and $\G \V M : \tau  @ A$ then
	$ M \in V^A $ or $M'$ exists such that $M \longrightarrow_s M'$.
\end{theorem}

>>>>

Substitution Lemma in \LMD\ is little more complicated than an ordinary one 
because there are six types of judgment and two types of substitution in \LMD.
Substitution Lemma on Terms states that term substitution $[z \mapsto P]$ preserves
typing, kinding, well-formed kinding, term equivalence, type equivalence, and king equivalence.
Substitution Lemma on Terms states the same for stage substitution $[\beta\mapsto\varepsilon]$.

@@

\begin{theorem}[Substitution Lemma on Stages]
	\begin{flalign*}
		\text{If\ } \G \V M:\tau @A
		&\text{\ then\ } \G[\beta \mapsto \varepsilon]\V M[\beta \mapsto \varepsilon]:\tau[\beta \mapsto \varepsilon] @A[\beta \mapsto \varepsilon].&\\
		\text{If\ } \G \V \tau::K @A
		&\text{\ then\ } \G[\beta \mapsto \varepsilon]\V \tau[\beta \mapsto \varepsilon]::K[\beta \mapsto \varepsilon] @A[\beta \mapsto \varepsilon].&\\
		\text{If\ } \G \V K\iskind @A
		&\text{\ then\ } \G[\beta \mapsto \varepsilon]\V K[\beta \mapsto \varepsilon] \iskind @A[\beta \mapsto \varepsilon].&\\
		\text{If\ } \G \V M\E N : \tau @A
		&\text{\ then\ } \G[\beta \mapsto \varepsilon]\V M[\beta \mapsto \varepsilon]\E N[\beta \mapsto \varepsilon] : \tau[\beta \mapsto \varepsilon]  @A[\beta \mapsto \varepsilon].&\\
		\text{If\ } \G \V \tau\E \sigma : K @A
		&\text{\ then\ } \G[\beta \mapsto \varepsilon]\V \tau[\beta \mapsto \varepsilon]\E \sigma[\beta \mapsto \varepsilon] : K[\beta \mapsto \varepsilon] @A[\beta \mapsto \varepsilon].&\\
		\text{If\ } \G \V K\E J @A
		&\text{\ then\ } \G[\beta \mapsto \varepsilon]\V K[\beta \mapsto \varepsilon]\E J[\beta \mapsto \varepsilon] @A[\beta \mapsto \varepsilon].&
	\end{flalign*}
\end{theorem}

@@

\begin{theorem}[Unique Decomposition]
	If $x:\tau @\varepsilon \notin \G$ and $\G \V M : \tau @ A$ then either
	\begin{enumerate}
		\item $ M \in V^A$, or
		\item there exist $B, E^A_B$, and $R^B$ such that $M = E^A_B[R^B]$ with $B = \varepsilon$ or $B = \beta$ for some $\beta$.
	\end{enumerate}
\end{theorem}

@@

\begin{theorem}[Progress]
	If $x:\tau @\varepsilon \notin \G$ and $\G \V M : \tau  @ A$ then
	$ M \in V^A $ or $M'$ exists such that $M \longrightarrow_s M'$.
\end{theorem}

### 2019-06-14 text/related-work.tex
分割コンパイルが失敗する
epsilon -> varepsilon.

In $\lambda^\TW$\cite{Tsukada}, Tsukada and Igarashi found modal logic which corresponds to a multi-stage calculus with environment classifiers and
show that run can be represented as application of $\epsilon$ to transition abstractions.
In \LTP\cite{Hanada2014}, Hanada and Igarashi extended $\lambda^\TW$ with CSP and discuss code residualization 
which allows us to dump the quoted code into an external file.

>>>>

In $\lambda^\TW$\cite{Tsukada}, Tsukada and Igarashi found modal logic which corresponds to a multi-stage calculus with environment classifiers and
show that run can be represented as application of $\varepsilon$ to transition abstractions.
In \LTP\cite{Hanada2014}, Hanada and Igarashi extended $\lambda^\TW$ with CSP and discuss code residualization 
which allows us to dump the quoted code into an external file.
